Glyph-maps for Visualizing Climate Data and Models

### 2011-09-28 paper.tex
fixed a couple of typos, added more to data explanation

Climate data is composed of measurements, such as temperature, precipitation, and winds, with a spatial and temporal context. The classic display for data of this type of is ``heatmap'', a tiled plot using color to display the value of the variable of interest as each location. When measurements are made at multiple time points it is common to display the data with small multiples, such as in Figure~\ref{fig:nasa-facet}. In this plot, a separate map is drawn for each month (columns) and year (rows), over a 6 year period of remotely sensed temperature data above Central America \citet{murrell:2010}. Color is used to display de-seasonalized temperature, with red mapped to high values and blue to low. The most noticeable feature is the strong red patch in the equatorial Pacific beginning mid-1997 and tapering out during 1998. This is the El Ni\~no event, a major temperature anomaly. More work is required to see more localized patterns, like the cooler land temperatures seen in early years.

@@

  \item GISTEMP surface temperature data provided by the Earth System Research
  Laboratory, Physical Sciences Division, National Oceanic and Atmospheric
  Administration. This data is provided on $2^{\circ}$ x $2^{\circ}$ grid over
  the entire globe, continental United States (XXX locations), measured
  monthly. Ground station data was de-seasonalized, differenced from from the
  1951-1980 temperature averages, and spatially averaged to obtain gridded
  measurements. We extracted the locations corresponding to the continental
  USA. Provided by the NOAA/OAR/ESRL PSD, Boulder, Colorado,
  USA, from their web site at \url{http://www.esrl.noaa.gov/psd/}.

>>>>

Climate data is composed of measurements, such as temperature, precipitation, and winds, with a spatial and temporal context. The classic display for data of this type of is ``heatmap'', a tiled plot using color to display the value of the variable of interest as each location. When measurements are made at multiple time points it is common to display the data with small multiples, such as in Figure~\ref{fig:nasa-facet}. In this plot, a separate map is drawn for each month (columns) and year (rows), over a 6 year period of remotely sensed temperature data above Central America \citep{murrell:2010}. Color is used to display de-seasonalized temperature, with red mapped to high values and blue to low. The most noticeable feature is the strong red patch in the equatorial Pacific beginning mid-1997 and tapering out during 1998. This is the El Ni\~no event, a major temperature anomaly. More work is required to see more localized patterns, like the cooler land temperatures seen in early years.

@@

\item GISTEMP surface temperature data provided on $2^{\circ}$ x
  $2^{\circ}$ grid over the entire globe, measured monthly. Ground
  station data was de-seasonalized, differenced from from the
  1951-1980 temperature averages, and spatially averaged to obtain
  gridded measurements. For the purposes of this paper, we extracted
  the locations corresponding to the continental USA. The data is made
  available by the Earth System Research Laboratory, Physical Sciences
  Division, National Oceanic and Atmospheric Administration, from
  their web site at \url{http://www.esrl.noaa.gov/psd/}.
